### 2014-11-11 tex/precise.tex

@@

In this section we present the Javalite rewrite rules for the
initialization of references in the symbolic heap summary. The
initialization rules are defined on the bi-partite graph consisting of
references and locations. In contrast, the heap summarization approach
creates a single branch that contains the summarization for all the
initialization in a single bi-partitate graph.

@@

The rewrite rule to initialize a heap in the symbolic summarization
technique is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns the
constraint location pair based on a lexical ordering of locations. The
init$_r$ method returns a new input reference $r_f$ while the
fresh$_f$ returns a new location of type $C$. The set $\rho$ contains
the set of reference location, $(r_a, l_a)$, pairs that the current
field may alias. The pair $(r_a, l_a)$ is contained in the set $ \rho
$ if the following conditions hold: (i) $r_a$ is part of the input
references initialized in the summary heap, (ii) $r_a$ is in the
pre-image of $l_a$, (iii) and the type of $l_a$ is the same as the
type of the field being initialized.

There are three constraint-location pair sets that correspond to
non-deterministic choices in GSE: (a) $\theta_\mathit{null}$ that
creates a constraint that states $r_f$ is null under which it points
to $l_\mathit{null}$, (b) $\theta_\mathit{new}$ the constraint that
$r_f$ is not null and it does not alias another reference in set
$\rho$ under which it points to the newly created location $l_f$, (c)
and $\theta_\mathit{alias}$ where for each element in the set $(r_a,
l_a) \in \rho$ it creates a constraints that $r_f$ is not null, points
to $r_a$, and does not point to any reference contained in
$(r_a^\prime, l_a^\prime) \in \rho$ such that $r_a \neq
r_a^\prime$. Additionally $\theta_\mathit{orig}$ allows for the
possibility that the field continues to remain uninitialized since the
heap is symbolic. We refer to this initialization in the summary
machine as conditional initialization. Finally, the fields of the
reference $r_f$ are marked as uninitialized.

In the post condition of the rule the reference $r_f$ points to
$\theta$ which contains the union of all the constraint location pairs
sets, and the field $f$ points to the new reference $r_f$ in the
summary heap.  Note that the summarize rule invoked repeatedly until
the set of unitialized constraint location pairs for field $f$ is
empty. In other words the set $\lambda$ is empty. This consititutes
the summarize end rule in~\figref{fig:symInit}.

\begin{figure*}[t]
\begin{center}
\begin{tabular}[c]{c|c|c|c}
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{origHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeXHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeYHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{l}
$\rho := \{ (r_1^i, r_1^i \neq r_\mathit{null}, l_1 \}$ \\
$\theta_\mathit{null} := \{ ( r_2^i = r_\mathit{null}, l_\mathit{null}) \}$\\
$\theta_\mathit{new} := \{ ( r_2^i \neq r_\mathit{null} \wedge r_2^i \neq r_1^i, l_2) \}$\\
$\theta_\mathit{alias} := \{ ( r_1^i \neq r_\mathit{null} \wedge r_2^i \neq r_\mathit{null} \wedge r_2^i = r_1^i, l_1) \}$\\
$\theta_\mathit{orig} := \{ \}$ \\
$\phi_{\mathit{1a}} := r_1^i = r_\mathit{null} $ \\
$\phi_{\mathit{1b}} := r_1^i \neq r_\mathit{null} $  \\
$\phi_{\mathit{2a}} := r_2^i = r_\mathit{null}$ \\
$\phi_{\mathit{2b}} := r_2^i \neq r_\mathit{null} \wedge r_2^i \neq r_1^i$ \\
$\phi_{\mathit{2c}} :=  r_2^i \neq r_\mathit{null} \wedge r_2^i = r_1^i $ \\
\end{tabular} \\
(a) & (b) & (c) & (d) \\
\end{tabular}
\end{center}
\caption{initialize this.x and this.y}
\label{fig:initHeap}
\end{figure*}

We visualize the initialization process in the symbolic summarization
technique in~\figref{fig:initHeap}. The heap
in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a local
reference. In~\figref{fig:initHeap} $r_0^s$ represents the reference
for the $\mathit{this}$ instance which has two fields $x$ and $y$ of
the same type. The reference $r_0^s$ points to the location
$l_0$. Note that when no constraint is specified, then there is a
implicit $\mathit{true}$ constraint. In~\figref{fig:initHeap} $r_0^s$
points to $l_0$ on $\mathit{true}$. The fields $x$ and $y$ point to
the uninitialized reference $r_\mathit{un}$. The reference
$r_\mathit{un}$ points to the uninitialized location $l_\mathit{un}$
on the $\mathit{true}$ constraint.

The graph in~\figref{fig:initHeap}(b) represents the summary heap
after the initialization of the $\mathit{this}.x$ field while the
graph in~\figref{fig:initHeap}(c) represents the summary heap after
the initialization of the $\mathit{this}.y$ field following the
initialization of $\mathit{this}.x$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the rewrite system when the initialization takes places. We also
define the constraints of references to their corresponding labels in
the graph.

@@

\subsection{Accessing a Field Reference}
\input{precise-fig-incomplete}

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for field
access and the other for field write. Let us consider the rewrite rule
for field access. The first check we perform is whether there exists a
constraint location pair for the $r$ being accessed such that the
location is not null and the constraint when conjuncted with the
global constraint is satisfiable. Next we extract all possible
constraints under which $r$ points to a null location such that the
constraint is satisfiable under the current global constraint,
$\phi_g$. The negation of these constraints are added to the global
constraint to create a new global constraint $\phi_g^\prime$. The
update to the global constraint ensures that access of the field $f$
happens only on non-null locations. The type $C$ of the field is
extracted and passed to rewrite rule that performs the initialization
of the symbolic heap. The initialization rules are described earlier
in this section.

Recall that during the initialization the rewrite rules
in~\figref{fig:symHeap} add input references that map to fields being
initialized.  Once the initialization is complete, however, we create
a new local reference $r^\prime$. An important property of the
references in the bi-partiate graph is that they are
\emph{immutable}. Hence we de-reference the initialized input
reference, assign its value to the the new local reference, and return
the local reference. In order to de-reference a field $r.f$ we define
a helper function which is called the value set.

@@

\subsection{Writing to a Field}

The reference $r_x$ is the base pointer whose field is being written
to while $r$ is the target reference. We look up the value of the base
reference in the environment $\eta(x)$. The set $\Psi_x$ is the set of
tuples of constraints, locations, and references which provide the
reference chains leading from $r_x$ to the reference of the field,
$r_\mathit{curr}$, being written to through $\phi$ and $l$.The goal is
to overwrite the $r_\mathit{curr}$ references with the target
references. Since the target of the write is $r$, we first check that the
location constraint pairs of $L(r)$ are satisfiable when accessed
through the $r_x$ chain. This is accomplished by the strengthing
function.

@@

Additionally, we also check for conditions where the write is not possible

\begin{comment}
\begin{figure}[t]
\begin{center}
\begin{tabular}[c]{l}
$\Psi_x = \{ (true, l_0, r_1^i) \}$\\
$ST (L, r_3^s, \phi, \phi_g)$ \\
$\theta = \{ (\phi_{2a}\; l_\mathit{null} ) (\phi_{2b}\; l_2) (\phi_{2c}\; l_1) \}$\\
$ST(L, r_0, \phi, \phi_g)$\\
$\theta = \{ \}$\\
\end{tabular}
\end{center}
\caption{field write for this.x = this.y sets}
\label{fig:faHeapSets}
\end{figure}
\end{comment}

>>>>

In this section we present the rules for the $\rsym$ relation. The are
three sets of rewrite rules: the initialization of symbolic
references, the field access and write rules, and finally checking the
equality and inequality of references. Here we describe each of the
rules. They consist the core of the symbolic heap summary approach.

@@

In this section we present the rewrite rules for the initialization of
references in the $\rsym$ relation for a symbolic heap summary. The
initialization rules are defined on the bi-partite graph consisting of
references and locations. In the $\rgse$ relation during
initialization of a symbolic references, it is non-deterministically
assigned to null, new instance of the object, and alias any previously
initialized symbolic reference. In contrast, the heap summarization
approach creates a single branch that contains the summarization for
all the initialization in a single bi-partitate graph.

@@

The rewrite rule to initialize a heap in the symbolic summarization
technique is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x, l_x)$ based on a lexical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $r_f$
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference location pairs, $(r_a, l_a)$, that the
field being initialized may alias. The pair $(r_a, l_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $r_a$ is
part of the input references initialized in the summary heap, (ii)
$r_a$ is in the pre-image of $l_a$, (iii) and the type of $l_a$ is the
same as the type of the field being initialized.

There are three constraint-location pair sets that correspond to
non-deterministic choices in GSE: (a) $\theta_\mathit{null}$ contains
a constraint $r_f = r_\mathit{null}$ paired with the location
$l_\mathit{null}$, (b) $\theta_\mathit{new}$ creates a pair with the
constraint $r_f$ is not null and it does not alias any reference in
set $\rho$ under and a newly created location $l_f$, (c) and
$\theta_\mathit{alias}$ where for each element in the set $(r_a, l_a)
\in \rho$ it creates a constraints that $r_f$ is not null, points to
$r_a$, and does not point to any reference contained in $(r_a^\prime,
l_a^\prime) \in \rho$ such that $r_a < r_a^\prime$ based on a lexical
ordering of initialized references. Additionally
$\theta_\mathit{orig}$ allows for the possibility that the field
continues to remain uninitialized since the heap is symbolic. We refer
to this initialization in the summary machine as conditional
initialization. Finally, the fields of the reference $r_f$ are marked
as uninitialized. In the post condition of the rule the reference
$r_f$ points to $\theta$ which contains the union of all the
constraint location pairs sets, and the field $f$ points to the new
reference $r_f$ in the summary heap.

\begin{comment}
 Note that the summarize rule invoked repeatedly until
the set of unitialized constraint location pairs for field $f$ is
empty. In other words the set $\Lambda$ is empty. This consititutes
the summarize end rule in~\figref{fig:symInit}.
\end{comment}

\begin{figure*}[t]
\begin{center}
\begin{tabular}[c]{c|c|c|c}
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{origHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeXHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeYHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{l}
$\rho := \{ (r_1^i, l_1 \}$ \\
$\theta_\mathit{null} := \{ ( r_2^i = r_\mathit{null}, l_\mathit{null}) \}$\\
$\theta_\mathit{new} := \{ ( r_2^i \neq r_\mathit{null} \wedge r_2^i \neq r_1^i, l_2) \}$\\
$\theta_\mathit{alias} := \{ ( r_1^i \neq r_\mathit{null} \wedge r_2^i \neq r_\mathit{null} \wedge r_2^i = r_1^i, l_1) \}$\\
$\theta_\mathit{orig} := \{ \}$ \\
$\phi_{\mathit{1a}} := r_1^i = r_\mathit{null} $ \\
$\phi_{\mathit{1b}} := r_1^i \neq r_\mathit{null} $  \\
$\phi_{\mathit{2a}} := r_2^i = r_\mathit{null}$ \\
$\phi_{\mathit{2b}} := r_2^i \neq r_\mathit{null} \wedge r_2^i \neq r_1^i$ \\
$\phi_{\mathit{2c}} :=  r_2^i \neq r_\mathit{null} \wedge r_2^i = r_1^i $ \\
\end{tabular} \\
(a) & (b) & (c) & (d) \\
\end{tabular}
\end{center}
\caption{initialize this.x and this.y}
\label{fig:initHeap}
\end{figure*}

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a local
reference. In~\figref{fig:initHeap} $r_0^s$ represents a local
reference for the $\mathit{this}$ reference which has two fields $x$
and $y$ of the same type; reference $r_0^s$ points to location
$l_0$. Note that when no constraint is specified in the graphs, then
there is a implicit $\mathit{true}$ constraint, for example, $r_0^s$
points to $l_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $r_\mathit{un}$.

% The reference $r_\mathit{un}$ points to the uninitialized location
%$l_\mathit{un}$ on the $\mathit{true}$ constraint.

The graph in~\figref{fig:initHeap}(b) represents the symbolic heap
summary after the initialization of the $\mathit{this}.x$ field while
the graph in~\figref{fig:initHeap}(c) represents the symbolic heap
summary after the initialization of the $\mathit{this}.y$ field
following the initialization of $\mathit{this}.x$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the summarize rule during the initialization of $\mathit{this}.y$. We
also define values of constraints used as labels in the graph.

@@

\subsection{Reading and Writing References}
\input{precise-fig-incomplete}

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). In both the field rewrite rules we first check we
perform is whether there exists a constraint location pair for the $r$
being accessed such that the location is not null and the constraint
when conjuncted with the global constraint is satisfiable. Next we
extract all possible constraints under which $r$ points to a null
location such that the constraint is satisfiable under the current
global constraint, $\phi_g$. The negation of these constraints are
added to the global constraint to create a new global constraint
$\phi_g^\prime$. The update to the global constraint ensures that
access or write of the field $f$ happens only on non-null
locations. There are rewrite rules for accessing and writing to fields
on null locations where it leads the current program execution state
to an error state. The rewrite rules for the accessing and writing to
null locations are presented in the supplemental document.

In the field-access rewrite rule in~\figref{fig:fHeap}(a) after the
non-nullness check we extract the The type $C$ of the field is
extracted and invoke the summarize rewrite rule
in~\figref{fig:symInit} that performs the initialization of the field
if it points to uninitialized locations. Once the initialization is
complete, we create a new local reference $r^\prime$. An important
property of the references in the bi-partiate graph is that they are
\emph{immutable}. Hence we create a new local reference and assign it
to the initialized reference, and return the local reference. In order
to de-reference a field $r.f$ we define a helper function which is
called the value set.

@@

The field-write rewrite rule in~\figref{fig:fHeap}(a) after the
non-nullness check we We look up the value of the base reference in
the environment $\eta(x)$.  The reference $r_x$ is the base pointer
whose field, $r_\mathit{curr}$ is being written to to while the
reference $r$ is the target reference. The set $\Psi_x$ contains
tuples $(\phi l r_\mathit{curr})$ of constraints, locations, and
references. These tuples represent access chains leading from $r_x$ to
the reference of the field, $r_\mathit{curr}$. The goal is to
overwrite the $r_\mathit{curr}$ references with the target
references. Since the target of the write is $r$, we first check that
the location constraint pairs of $L(r)$ are satisfiable when accessed
through the $r_x$ chain. This is accomplished by the strengthening
function.

@@

Note that the write is conditional. In the case that $\phi$ is true
then $r_\mathit{curr}$ will point to the constraint location pairs of
$L(r)$ while if $\phi$ is false then $r_\mathit{curr}$ will continue
to point to the constraint location pair it is currently pointing to.
Again since the references are immutable we create a new reference for
each $r_\mathit{curr}$ and point them to the target constraint
location pairs.

Consider the example shown in~\figref{fig:fHeap}(b) where we assign
$\mathit{this}.x := \mathit{this}.y$. Note that
in~\figref{fig:initHeap}(c) $\mathit{this}.x$ is represented by
$r_1^i$. After applying the field-write rule the reference $r_1^i$ is
replaced by the fresh reference $r_4$ which now points to target
$\mathit{this}.y$ represented by $L(r_2^i)$.

\begin{comment}
\begin{figure}[t]
\begin{center}
\begin{tabular}[c]{l}
$\Psi_x = \{ (true, l_0, r_1^i) \}$\\
$ST (L, r_3^s, \phi, \phi_g)$ \\
$\theta = \{ (\phi_{2a}\; l_\mathit{null} ) (\phi_{2b}\; l_2) (\phi_{2c}\; l_1) \}$\\
$ST(L, r_0, \phi, \phi_g)$\\
$\theta = \{ \}$\\
\end{tabular}
\end{center}
\caption{FIXME: When will I get a caption}
\label{fig:faHeapSets}
\end{figure}
\end{comment}
